SoK: Defending Against Web Application Attacks: From Research Approaches to Practical Tools

### 2014-11-14 docs/questioning.tex
more changes / corrections

\noindent
Attackers that would input the string {\tt anything' OR 'x'='x},
could obtain the administrator's password by email.
Savvy programmers can use certain API functions, such as {\sc php}'s
{\tt mysql\_real\_escape\_string()} to detect malformed input, or
better use prepared {\sc sql} statements, instead of statement
templates. Unfortunately,
the increasing number of {\sc sql} injection attacks
suggest that programmers are not always that careful.
Using similar techniques, malicious users can mount other exploits based on
{\sc dsl}s such as {\sc xp}ath~\cite{SW06,CDL07,MKS09},
{\sc xml}~\cite{MSM13} and {\sc json}~\cite{SMS13}. With this kind of
attacks, a malicious user can view sensitive information, destroy or
modify protected data, or even crash the entire application.

@@

JavaScript injection attacks make up a large subset of dynamic
language code injection attacks. Such attacks are enabled when a web
application accepts and redisplays data of uncertain origin without
proper validation and filtering. Based on this flaw, an attacker
can manage to inject a script in the JavaScript engine of a browser
and alter its execution flow~\cite{ELX07}. For instance,
consider a web page that prints the value
of a query parameter ({\tt query}) from the
page's {\sc url} as part of the page's content
without escaping the value. Attackers
can take advantage of this and inject an {\tt iframe} tag
into the page to steal a user's cookie and
send it via an image request to a web site
under their control (malicious.com).
This could be achieved by including the following
link to the malicious web site (or send it via phishing
email) and induce the user to click on it:

@@

\noindent
JavaScript injection is considered a critical issue in web application security
mainly because it is associated with major vulnerabilities such as
{\sc xss} attacks~\cite{SG07} and Cross-Channel Scripting
({\sc xcs}) attacks~\cite{W10,BBB09}. In many cases {\sc xss} attacks
involve the injection of both {\sc html} and JavaScript code.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps-new.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model for dynamic and domain specific
language code injection in web applications.
Transitions are labeled with the different attacks that use that path,
while the numbers next to attack labels denote the 
sequence of steps for a particular attack.
Points where different types of defenses
detect or prevent attacks are marked with an `X' symbol:
{\sc ub} (browser):~\cite{KJKV09,LV09,TNH07,NSS06,APKLM10,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,VFJKKV07,SLMS14,BV08},
{\sc s}t{\sc b} (en route from the server to the browser):~\cite{RDWDE07,JKK06a,GC09,JB07,NLC07,WPLKK09,JEP08,PS11},
{\sc e}i{\sc q} (embedded in a query en route to the database):~\cite{BWS05,SW06,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point i{\sc db} (at the database):~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

Detection mechanisms vary on the location where they detect an attack.
There are four different points where an attack can be caught, as seen
in Figure~\ref{fig:attacks}:
\begin{itemize}
\item At the user's browser (Point {\sc ub}).
\item En route from the server to the user's browser---in most cases
within a proxy (Point {\sc s}t{\sc b}).
\item Embedded in a query, but before it reaches the server's database
  (Point {\sc e}i{\sc q}).
\item After the malicious code reaches the server's database
  (Point i{\sc db}).
\end{itemize}

@@

In the case of {\sc dsl} injection attacks, mechanisms check 
the query before the inclusion of user input with the one
resulting after the inclusion of user input.
Both mechanisms that implement this approach,
{\sc sqlg}uard~\cite{BWS05} and
{\sc sql}check~\cite{SW06} are pretty similar
and detect the attack before a query reaches the
database (Point {\sc e}i{\sc q}).
Contrary to {\sc sqlg}uard and other
mechanisms, {\sc sql}check has been
extensively tested in terms of accuracy
as can be seen in Table~\ref{tab:comp}.
A disadvantage of those mechanisms is that the
application must be modified in every code fragment
that sends an {\sc sql} query to the database for execution.

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
an exception is raised (Point {\sc e}i{\sc q}).

It is possible to apply further checks when it is established that
tainted data have reached a sink (Point {\sc e}i{\sc q}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Point {\sc e}i{\sc q}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of the  {\sc php} array data structure.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (Point {\sc s}t{\sc b}).

@@

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e., {\sc sql} or {\sc
  xp}ath queries) in various forms, and detect attacks as outliers
from the set of valid code statements. An early approach, {\it {\sc
    didafit}}~\cite{LLW02} detects {\sc sql} injection (Point
{\sc sdb}) attacks by recording all database transactions
stripped from user input. Subsequent
refinements by Valeur et al.~\cite{VMV05} tagged each transaction with
the corresponding application as an extension of their anomaly
detection framework called {\it libAnomaly}. {\it {\sc
    sd}river}~\cite{MS09,MKS09,MKLS11} is a signature-based mechanism
that prevents {\sc sql} and {\sc xp}ath injection attacks. The
signatures generated during a training phase are based on features
that can depend either on the code statement or on its execution
environment (i.e., the stack trace). Then, at runtime, the mechanism
checks all statements for compliance and can block code statements
containing injected elements (Point {\sc e}i{\sc q}). {\sc
  amnesia}~\cite{HO05,HO06,HO05b} is a tool that also detects {\sc
  sql} injection attacks (Point {\sc e}i{\sc q}) by associating a query model
with the location of every {\sc sql} statement within the application.
Then, at runtime, it monitors the application's execution to detect
when {\sc sql} statements diverge from the expected model.

@@

{\it Diglossia}~\cite{SMS13} combines positive taint tracking together
with parse-tree validation. Diglossia was based on the theory of Ray
and Ligatti~\cite{RL12b} (which we saw in Section~\ref{sec:attacks})
to detect {\sc dsl} code injection attacks. In addition, it is
actually the first, and so far the only framework that detects {\sc
  json} injection attacks. When an application computes an output
string (query), Diglossia computes a ``shadow'' of this string.
Specifically, it maps all characters introduced by the application to
a shadow character set. This set does not contain any characters
coming from the tainted input. Then, the scheme creates the tree of
the query that is about to be executed and compares it with the parsed
tree of the ``shadow'' (Point {\sc e}i{\sc q}). If the trees are not
equal the application is probably under attack. Note that Diglossia
can be bypassed in the same way as other taint tracking
approaches~\cite{NBR14}.

>>>>

\noindent
Attackers that would input the string {\tt anything' OR 'x'='x},
could obtain the administrator's password by email.
Savvy programmers can use certain {\sc api} functions, such as {\sc php}'s
{\tt mysql\_real\_escape\_string()} to detect malformed input, or
better use prepared {\sc sql} statements, instead of statement
templates. Unfortunately,
the increasing number of {\sc sql} injection attacks
suggest that programmers are not always that careful.
Using similar techniques, malicious users can mount other exploits based on
{\sc dsl}s such as {\sc xp}ath~\cite{SW06,CDL07,MKS09},
{\sc xml}~\cite{MSM13} and {\sc json}~\cite{SMS13}. With this kind of
attacks, a malicious user can view sensitive information, destroy or
modify protected data, or even crash the entire application.

@@

JavaScript injection attacks make up a large subset of dynamic
language code injection attacks. Such attacks are enabled when
a web application accepts
and redisplays data of uncertain origin without
proper validation and filtering. Based on this flaw, an attacker
can manage to inject a script in the JavaScript engine of a browser
and alter its execution flow~\cite{ELX07}.
JavaScript injection attacks are considered a critical issue
in web application security mainly because it is associated with
major vulnerabilities such as {\sc xss} attacks~\cite{SG07} and
Cross-Channel Scripting ({\sc xcs}) attacks~\cite{W10,BBB09}.
A typical {\sc xss} example that involves JavaScript injection
is the following: consider a web page that prints the value
of a query parameter ({\tt query}) from the
page's {\sc url} as part of the page's content
without escaping the value. Attackers
can take advantage of this and inject an {\tt iframe} tag
into the page to steal a user's cookie and
send it via an image request to a web site
under their control (malicious.com).
This could be achieved by including the following
link to the malicious web site (or send it via phishing
email) and induce the user to click on it:

@@

\noindent
Note that in many cases {\sc xss} attacks
involve the injection of both {\sc html} and JavaScript code.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps-new.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model for dynamic and domain specific
language code injection in web applications.
Transitions are labeled with the different attacks that use that path,
while the numbers next to attack labels denote the 
sequence of steps for a particular attack.
Points where different types of defenses
detect or prevent attacks are marked with an `X' symbol:
{\sc ub} (at the browser):~\cite{KJKV09,LV09,TNH07,NSS06,APKLM10,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,VFJKKV07,SLMS14,BV08},
{\sc s}t{\sc b} (en route from the server to the browser):~\cite{RDWDE07,JKK06a,GC09,JB07,NLC07,WPLKK09,JEP08,PS11},
{\sc dbal} (at the database abstraction layer):~\cite{BWS05,SW06,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
i{\sc db} (at the database):~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

Detection mechanisms vary on the location where they detect an attack.
There are four different points where an attack can be caught, as seen
in Figure~\ref{fig:attacks}:
\begin{itemize}
\item At the user's browser (Point {\sc ub}).
\item En route from the server to the user's browser---in most cases
within a proxy (Point {\sc s}t{\sc b}).
\item At the database abstraction layer, before it reaches the server's database
  (Point {\sc dbal}).
\item After the malicious code reaches the server's database
  (Point i{\sc db}).
\end{itemize}

@@

In the case of {\sc dsl} injection attacks, mechanisms check 
the query before the inclusion of user input with the one
resulting after the inclusion of user input.
Both mechanisms that implement this approach,
{\sc sqlg}uard~\cite{BWS05} and
{\sc sql}check~\cite{SW06} are pretty similar
and detect the attack before a query reaches the
database (Point {\sc dbal}).
Contrary to {\sc sqlg}uard and other
mechanisms, {\sc sql}check has been
extensively tested in terms of accuracy
as can be seen in Table~\ref{tab:comp}.
A disadvantage of those mechanisms is that the
application must be modified in every code fragment
that sends an {\sc sql} query to the database for execution.

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
an exception is raised (Point {\sc dbal}).

It is possible to apply further checks when it is established that
tainted data have reached a sink (Point {\sc dbal}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Point {\sc dbal}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of the  {\sc php} array data structure.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (Point {\sc s}t{\sc b}).

@@

Training methods that detect {\sc dsl}-driven injection attacks
generate and store valid code statements (i.e., {\sc sql} or {\sc
  xp}ath queries) in various forms, and detect attacks as outliers
from the set of valid code statements. An early approach, {\it {\sc
    didafit}}~\cite{LLW02} detects {\sc sql} injection (Point
{\sc sdb}) attacks by recording all database transactions
stripped from user input. Subsequent
refinements by Valeur et al.~\cite{VMV05} tagged each transaction with
the corresponding application as an extension of their anomaly
detection framework called {\it libAnomaly}. {\it {\sc
    sd}river}~\cite{MS09,MKS09,MKLS11} is a signature-based mechanism
that prevents {\sc sql} and {\sc xp}ath injection attacks. The
signatures generated during a training phase are based on features
that can depend either on the code statement or on its execution
environment (i.e., the stack trace). Then, at runtime, the mechanism
checks all statements for compliance and can block code statements
containing injected elements (Point {\sc dbal}). {\sc
  amnesia}~\cite{HO05,HO06,HO05b} is a tool that also detects {\sc
  sql} injection attacks (Point {\sc dbal}) by associating a query model
with the location of every {\sc sql} statement within the application.
Then, at runtime, it monitors the application's execution to detect
when {\sc sql} statements diverge from the expected model.

@@

{\it Diglossia}~\cite{SMS13} combines positive taint tracking together
with parse-tree validation. Diglossia was based on the theory of Ray
and Ligatti~\cite{RL12b} (which we saw in Section~\ref{sec:attacks})
to detect {\sc dsl} code injection attacks. In addition, it is
actually the first, and so far the only framework that detects {\sc
  json} injection attacks. When an application computes an output
string (query), Diglossia computes a ``shadow'' of this string.
Specifically, it maps all characters introduced by the application to
a shadow character set. This set does not contain any characters
coming from the tainted input. Then, the scheme creates the tree of
the query that is about to be executed and compares it with the parsed
tree of the ``shadow'' (Point {\sc dbal}). If the trees are not
equal the application is probably under attack. Note that Diglossia
can be bypassed in the same way as other taint tracking
approaches~\cite{NBR14}.
